### 2019-07-21 paper_round2.Rmd
changes following revision from PCI

---
title: Niche complementarity between pollinators increases plant reproductive success 
author: "Magrach, A, Molina, FP, Bartomeus, I"
csl: ecology-letters.csl
output:
  word_document: default
  html_document: default
  pdf_document: default
editor options:
  chunk_output_type: console
bibliography: library.bib
---

**Niche complementarity between pollinators increases plant reproductive success**

Ainhoa Magrach^1^, Francisco P. Molina^2^, Ignasi Bartomeus^2^

Short running title: Network structure effects on ecosystem functioning

^1^Basque Centre for Climate Change-BC3, Edif. Sede 1, 1º, Parque Científico UPV-EHU, Barrio Sarriena s/n, 48940, Leioa, Spain

^2Estación Biológica de Doñana (EBD-CSIC), Avda. Américo Vespucio 26, Isla de la Cartuja, 41092, Sevilla, Spain

Type of article: Letter

Words in Abstract: 148 words

Words in main text (excluding abstract, acknowledgements, references, table and figure legends): 4,451

Number of references: 48

Number of figures, tables and text boxes: 7

Corresponding author: ainhoa.magrach@bc3research.org, nacho.bartomeus@gmail.com

Keywords: nestedness, niche complementary, fruitset, pollination, plant-pollinator interactions

\nextpage

**Abstract**

Declines in pollinator diversity and abundance have been reported across different regions, with implications for the reproductive success of plant species. However, research has focused primarily on pairwise plant-pollinator interactions, largely overlooking community-level dynamics. Yet species do not interact in isolation, they are embedded within larger networks whose structure can affect pollinator functional roles and, ultimately, the pollination services they deliver to plants. Here, we present one of the first efforts linking pollinator visitation to plant reproduction from a community-wide perspective using a well-replicated dataset encompassing 16 well-resolved plant-pollinator networks and data on reproductive success for 19 plant species from Mediterranean shrub ecosystems. We find that, for prediction purposes, information on simple visitation metrics is sufficient. Contrastingly, a mechanistic understanding of the pathways through which differences in pollinator diversity translate into changes in reproductive success, requires additional information on network structure, particularly that reflecting niche complementarity between pollinators. Specifically, we find a positive effect of increasing niche complementarity between pollinators on plant reproductive success.  
\nextpage

**Introduction**

Pollinators provide key services to plants by facilitating pollen flow between individuals. The recent declining trends found for some pollinator species in some regions of the planet ([@Potts2010; @I.2019]) have led many researchers to focus on the functional impacts of these changes in pollinator diversity, with a major focus being placed on the consequences for plant reproductive success [@Biesmeijer2006].

  Many research efforts have targeted the reproductive success of individual plant species [@Albrecht2012a; @doi:10.1111/1365-2745.13055], and used relatively simple visitation metrics (e.g., the number of pollinator species that visit a plant or the number of visits they perform) to explain the differences observed across different plant individuals. Contrastingly, community-level analyses remain scarce [@Bennett2018]. Yet plants and pollinators do not interact in isolation, but rather are embedded within larger networks of interactions encompassing other plant and pollinator species. We are thus missing an important part of the picture, which includes the direct interactions between the whole ensemble of plants and pollinators, but also the indirect interactions between species within one guild (e.g., plants) through their shared resources [@Carvalheiro2014; @Lazaro2014; @Mayfield2017; @Pauw2013; @Johnson2019]. Understanding how changes in pollinator diversity and interaction structure affect whole community assemblages is thus a major challenge that requires attention.

  The few pollination studies that have analysed the effects of changing pollinator diversity for reproductive success at the community level have done so using mainly experimental setups. As an example, a study that experimentally recreated a plant community with 9 plant species and differing levels of pollinator diversity across different enclosures, found that not only pollinator species diversity had an effect for average reproductive success, but also they found that network structure had an important effect [@Frund2013]. In particular, these authors found that niche complementarity between pollinators, in terms of plant species and temperature coverage (a measure of the overlap in the use of plant resources and optimum temperature activity) had a positive effect for average seed set at the community level [@Frund2013]. This provides added information on the effects of biodiversity for ecosystem functioning, suggesting that not only the diversity of species present, but also the diversity of roles and ways in which a community is structured, are determinant factors.

  Indeed, theoretical research has long suggested that the structure of multitrophic communities has an effect for ecosystem functioning (reviewed in [@Thompson2012]). This line of research, primarily driven by food-web studies, has greatly advanced theory, but these ideas have not yet been tested using empirical data (but see [@Poisot2013]). Specifically, a major knowledge gap resides in understanding which aspects of structure determine which aspects of function [@Thompson2012]. We are now at a point in which there is considerable understanding on the attributes characterizing mutualistic interaction networks [@Bascompte2007b] and we have a substantial understanding of how these attributes vary along different types of gradients [@Pellissier2017; @Tylianakis2016]. Especially, in the case of pollination, we have ample knowledge on the attributes that shape these mutualistic interactions at the community level. Amongst them is the prevalence of nested structures, i.e., arrangements where specialist species interact with a subset of the species that generalists interact with [@Bascompte2003] and which is thought to promote species diversity [@Bastolla2009]; or the presence of asymmetric specialization, i.e., specialists tend to interact with generalist species as a pervasive feature [@Vazquez2004]. However, the pathways by which the structure of these networks of interactions affect key ecosystem functions such as plant reproduction remain to be understood [@Winfree2013]. The time is thus ripe to use the existing knowledge around plant-pollinator network structures to explore the relationship between network structure and ecosystem functioning, with special emphasis being placed on the underlying mechanisms that drive these relationships.

  Here, we present one of the first efforts linking pollinator visitation and plant reproductive success at the community level using empirical data on plant-pollinator interaction networks and plant reproductive success. To this end, we use a well-replicated dataset encompassing 16 well-resolved plant-pollinator interaction networks coupled with data on the reproductive success of 19 plant species recorded in Mediterranean shrub ecosystems within the area of influence of Doñana National Park in SW Spain, i.e., within the limits of the Natural Space of Doñana as defined by the local government (Junta de Andalucía). Our study focuses on understanding whether adding information on network structure to previously used simple visitation metrics (e.g., the number and diversity of pollinator species visiting a plant species) aids in better explaining the differences observed in community-wide reproductive success. In doing so, we conducted our analyses focusing on reproductive success at two different levels: (i) at the species level by considering the effect of the position of a focal species within the larger network and its impact on its individual reproductive success, and (ii) at the site level, by evaluating how attributes that describe the whole site might affect average values of reproductive success. Specifically, our study focuses on the how the interplay between niche complementarity and redundancy determines reproductive success. Plant reproductive success requires of the delivery of conspecific pollen and thus of a certain degree of niche complementarity. Yet, greater values of nestedness, which imply redundancy in species functions, are thought to promote species diversity [@Bastolla2009] and stability [@Thebault2010a] within plant-polllinator networks. At present, we do not know how either of these network characteristics affects the functions performed by pollinators. Finally, in addition to average values, we also evaluate whether network structure helps explain differences in equity in reproductive success across species within a community, as a measure of evenness in the pollination service delivered.

  Our results suggest that for prediction purposes, information on simple visitation metrics, particularly regarding the pollinator richness visiting a plant species, is sufficient. However, we find that a mechanistic understanding of the pathways through which differences in pollinator diversity translate into changes in reproductive success requires additional information on network structure, notably information on the complementarity between the functions performed and the niches occupied by different pollinator species.

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

#LOAD ALL NECESSARY DATA AND PACKAGES

#R PACKAGES

library(devtools)
#Then install the data package. Reprot issues 
#install_github("ibartomeus/BeeFunData", ref = "fun")

library(BeeFunData)#this package holds all the necessary data
library(plyr)
library(dplyr)
library(ggplot2)
library(grid)
library(iNEXT)
require(Hmisc)
library(knitr)
library(lme4)
library(lmerTest) #to calculate coeff and p values
library(car) #to calculate VIF
library(bbmle)  #to select best model based on AIC
library(ggplot2)
library(DHARMa) #to check for problems in residuals
library(reshape2)
library(spaa)
library(cowplot)
#devtools::install_github("hohenstein/remef")
library(remef)
library(glmmTMB)
library(vegan)
library(piecewiseSEM)
library(MuMIn)

#LOAD ALL DATA

d<-all_interactions
f<-fruitset
s<-seedset

f$Plant.sp<-paste(f$Plant_genus, f$Plant_species)
f$Plant.sp<-as.factor(f$Plant.sp)

s$Plant.sp<-paste(s$Plant_genus, s$Plant_species)
s$Plant.sp<-as.factor(s$Plant.sp)

#1st eliminate frutos perdidos
wordList <- c("fruto perdido","frutos perdidos")

f2<-subset(f, Notes %nin% wordList)

#calculate column with fruit set as
f2$tot.fruitset<-(f2$Fruit_Yes +  f2$Fruit_preyed + f2$Fruit_dispersed)/(f2$Fruit_Yes +f2$Fruit_No + f2$Fruit_preyed + f2$Fruit_dispersed)

s2<-subset(s, Notes %nin% wordList)

#calculate average seed set per plant individual, there are several fruit measurements per plant and would lead to 3 nested random factors

s3 <- ddply(s2, c("Site_ID", "Plant.sp", "Plant.ID"), summarise,
                         mean.seedn    = mean(Seeds2),
                          mean.seedw = mean(Seed.weight),
                          mean.fruitweight = mean(Fruit.weight2))

###### read network and species level files (network analyses to be found at "...scripts/network analysis.R")

outsp.pl.site<-read.csv("SITE_plant_species_level_metrics.csv")
outsp.site<-read.csv("SITE_species_level_metrics.csv")
out.site<-read.csv("SITE_network_level_metrics.csv")

outsp.pl.site.sinout<-read.csv("SITE_plant_species_level_metrics_sinout.csv")
outsp.site.sinout<-read.csv("SITE_species_level_metrics_sinout.csv")
out.site.sinout<-read.csv("SITE_network_level_metrics_sinout.csv")

#check number of sps per site surveyed in the focal analyses

foc<-subset(d, d$Out=="focal")
foc2<-ddply(foc, c("Site_ID", "Plant_gen_sp"), summarise, 
              tot.species = length(Plant_gen_sp)) #number of species per site
foc3<-ddply(foc2, c("Site_ID"), summarise, 
            tot.species = length(Site_ID)) #number of species per site

```

**Methods**

_Plant pollinator interactions_

Our study was conducted in SW Spain within the area of influence of Doñana National Park (Fig. 1). Here, we surveyed 16 Mediterranean woodland patches with an average distance of 7 km between them (min= 3 km, max= 46.5 km). Each site was surveyed 7 times during the flowering season of 2015 (from February to May) following a 100-m x 2 m transect for 30 mins. Along each transect, we identified all plant species and recorded all the floral visitors that landed on their flowers during each 30-min period. Only floral visitors (from now on referred to as pollinators) that could not be identified in the field were captured, stored and identified in the laboratory by FPM with the aid of experts in the different taxonomic groups (see acknowledgements). In addition, at each round we conducted 3 minutes of focal observations recording all floral visitors observed on 3 plant individuals per species belonging to the 19 most common (based on previous surveys) plant species across the study area (mean $\pm$ SD: `r round(mean(foc3$tot.species), digits=2)` $\pm$ `r round(sd(foc3$tot.species), digits=2)` species per site). Furthermore, we included some interactions between plant and pollinator individuals that were not observed during the sampling but that were opportunistically recorded immediately before or after the sampling periods, as some of these interactions are difficult to document and might be important to define network structure [@Jordano2016]. These opportunistic interactions represented `r round(table(d$Out)[2]/(table(d$Out)[1]+table(d$Out)[2]+table(d$Out)[3])*100, digits=2)` of all interactions recorded. All surveys were done under similar weather conditions, avoiding windy or rainy days. Surveys were done during mornings and afternoons with the sampling order being established randomly.

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

f.inds<-ddply(f, c("Site_ID", "Plant.sp"), summarise, 
                   tot.inds = length(Plant.sp)) #number of individuals per plant sps

tableS1<-ddply(f.inds, c("Site_ID"), summarise, 
                   tot.sps = length(Site_ID)) #number of plant sps per site

s.fruitnum<-ddply(s, c("Site_ID", "Plant.sp"), summarise, 
                   tot.fruits = length(Plant.ID)) #number of fruits per individual
```

_Plant reproductive success_

Within each site, we marked between 3 and `r max(f.inds$tot.inds)` individuals (mean $\pm$ SD: `r round(mean(f.inds$tot.inds), digits=2)` $\pm$ `r round(sd(f.inds$tot.inds), digits=2)`) belonging to `r min(tableS1$tot.sps)` to `r max(tableS1$tot.sps)` plant species (mean $\pm$ SD:`r round(mean(tableS1$tot.sps), digits=2)` $\pm$ `r round(sd(tableS1$tot.sps), digits=2)`), depending on the availability and presence of flowers during the sampling events. For each individual, at the end of the season, we 
recorded fruit set (i.e. the proportion of flowers that set fruit), the average number of seeds per fruit and the average fruit and seed weight per fruit (`r min(s.fruitnum$tot.fruits)`-`r max(s.fruitnum$tot.fruits)` fruits subsampled, mean $\pm$ SD: `r round(mean(s.fruitnum$tot.fruits), digits=2)` $\pm$ `r round(sd(s.fruitnum$tot.fruits), digits=2)`). Our survey included a total of `r nlevels(f2$Plant.sp)` different plant species across our `r nlevels(f2$Site_ID)` sites. Plants species were selected based on their availability with sampling being focused on the most abundant plant species. The values at the species level were then averaged per site to calculate unique reproductive success measures at the site level.

_Data analyses_

In order to evaluate the completeness of our sampling of the pollinator and plant community as well as that of their interactions, we estimated the asymptotic number of species present, a non-parametric estimator of species richness for abundance data. This estimator includes non-detected species, and allowed us to calculate the proportion detected with our original sampling data. We used Chao 1 asymptotic species richness estimators [@Chao2009] and estimated the richness of pollinators, plants and plant–pollinator links accumulated as sampling effort increased up to 100% sampling coverage using package iNEXT [@Hsieh2016] within the R environment [@RDevelopmentCoreTeam2011]. We then extracted the values covered by our sampling.

  In order to analyse how differences in network structure might affect plant reproductive success, we constructed weighted plant-pollinator interaction networks by pooling the data for the 7 rounds of sampling. We thus obtained one weighted interaction network per site, representing the frequency of visits by different pollinator species to different plant species. For each network, we then proceeded to extract a series of relevant network metrics at the species and site levels.

_Species-level network analysis_

At the species level, we focused on attributes defining the position of a focal plant species within the larger community. As such, we considered two metrics providing complementary non-redundant information: (i) average niche overlap in terms of pollinators between a focal plant species and each of the other plant species in the community, which estimates the potential indirect interactions between different plant species through shared resources (in this case pollinators), and (ii) centrality, which depicts the importance of the role played by a plant species within the larger community (as resource for a large number of pollinator species) and its contribution to network cohesiveness. 
 Niche overlap was calculated as the average overlap in pollinator species visiting a focal plant and each of the other plants in the community using the Morisita overlap index, a measure of similarity between two sets of data [@Zhang2016]. As a measure of centrality we used weighted closeness centrality, which represents the number of shortest paths going through a focal plant based on a unipartite projection of the bipartite plant-pollinator network using a weighted representation of the network [@Dormann2009]. Here, links between plant species represent shared pollinator species.

_Site-level network analysis_

At the site level, we followed the same logic as the one presented at the species level. Thus, we also calculated two network metrics providing complementary non-redundant information. In this case we focused on (i) nestedness and (ii) niche complementarity.

  Nestedness is the property by which specialists interact with a subset of the species that generalists interact with [@Bascompte2003]. Although there is an ongoing debate in the literature, some studies have found that nested networks are more stable and resilient to perturbations because nestedness promotes a greater diversity by minimizing competition among species in a community [@Bastolla2009]. However, many network attributes vary with network size and complexity [@Bluthgen2006]. In the case of nestedness, we know it can be affected by network size and connectance [@Song2017]. An approach that is often used to correct for this, is to use null models and to compare null-model corrected nestedness values across different networks. However, this approach has been recently shown to present the same issues, as z-scores also change with network size and connectance [@Song2017]. We thus followed the advice provided by Song et al. (2017) by using a normalized value of the widely-used nestedness metric NODF [@Almeida-Neto2011], $NODF_c$. This normalized value is calculated as $NODF_c = NODF_n/(C * log(S))$, where C is connectance and S is network size. $NODF_n$ is calculated as $NODF/max(NODF)$, which is independent of network size and thus comparable across different networks [@Song2017].

  Niche complementarity metrics are important because plant reproductive success depends on the delivery of conspecific pollen and thus of a certain level of specialization or niche divergence (reviewed in [@Brosi2016]). To calculate niche complementarity, we used a community-level measure defined as the total branch length of a functional dendrogram based on qualitative differences measured using a Euclidean distance in visitor assemblages between plants [@Devoto2012; @Petchey2007a]. All network metrics were calculated using package bipartite [@Dormann2009].

  All of these metrics were calculated using all the data as well as for the subset of the data excluding interactions observed outside of sampling periods. Differences between results are minimal for both and thus we will only present results for the analysis using the full dataset.

_Statistical analyses_

In order to evaluate whether adding information on network structure improves our ability to explain differences in reproductive success - both at the species and the site level - we used generalized linear (GLMs) and generalized linear mixed models (GLMMs). In both cases (species and site-level models) we fit two types of models: (i) model 1, that only included simple visitation metrics and (ii) model 2 that additionally included information on network structure.

  At the species level, response variables included the fruit set for different individuals of each species, analyzed using a binomial distribution, the average number of seeds per fruit, analyzed using a normal distribution, and the average values of fruit and seed weight fitted to Poisson distributions. The number of seeds per fruit was centered and scaled (i.e., we subtracted column means and divided by standard deviation) to allow meaningful comparisons across species with contrasting life histories. As explanatory variables, model 1 included pollinator richness, and the total number of visits received by that plant species; while model 2 added the two network attributes calculated at the species level: average niche overlap and centrality. For both models, we included plant species nested within site as a random effect to account for the non-independence of several individuals measured for the same plant species within each site.

	At the site level, we upscaled our species-level analyses. As response variable we had the average reproductive success per site (i.e., average fruit set analyzed using a binomial distribution, average number of seeds per fruit and average fruit and seed weight using a normal distribution). We thus had a single value per site and no random effects are needed in this case. In this case, model 1 included total pollinator richness and total pollinator abundance (i.e. number of visits received by all plants within the community) as explanatory variables. Model 2, in turn, added information on network structure by including nestedness and niche complementarity as explanatory variables.

  Average values of reproductive success at the site level can be driven to a large extent by a single plant species. Yet, what will determine the persistence of a diverse plant community, is the presence of some sort of “equity” or evenness in reproductive success across the whole community. We therefore calculated a measure of equity in reproductive success at the site level as the proportion of species with normalized (between 0 and 1) average fruit set values that were above 0.5. We repeated this using 0.25 and 0.75 thresholds. We then used the same framework as that used for species and site-level analyses and fit the same models 1 and 2 GLMs, but using equity in reproductive success as the response variable fitting a binomial distribution. In all cases, we used variance inflation factors to check for collinearity between explanatory variables. Additionally, we ran residual diagnostics to check if model assumptions were met. Then, we used information criterion methods to find the best model and selected the ones with the lowest Akaike Information Criterion (AIC) value. Whenever the difference between the AIC of both models was < 2 ($\Delta AIC<2$), we considered that both models were equally good [@Burnham2011]. All predictor variables were centered prior to analysis.

  Finally, we tested whether the importance of network structure in explaining differences in equity in reproductive success within communities increases with the number of plant species being considered. We expect that when only one plant species is considered, then the importance of network structure will be negligible, while we expect this importance to increase as more plant species are considered (up to a maximum number of 6 species which is the maximum we have measured in our study at a particular site).

  To test this, we ran a simple simulation in which the number of species considered increased at each step and for each step we re-calculated equity in reproductive success. Instead of drawing plant species randomly for each step, we tested all possible combinations for each plant number level and network, as the number of combinations is small (e.g. for n = 3 plant selected out of 6 there is only 20 possible combinations). Then, we tested if the relationship between equity in reproductive success and functional complementarity (given its importance in determining differences in reproductive success, see Results section) changes as a function of the number of plants considered within our simulated communities. To this end, for each level of species number considered, we randomly selected one of the generated equity values across each of the 16 communities, and regressed these 16 values against our network level predictor and extracted the model estimates. We repeated this process 1,000 times and averaged all estimates. We expect that the more plants considered, the larger the resulting average estimates will be. Note that we only interpret the mean effects, as the variance among different plant number of species considered depends on the initial number of combinations.

**Results**

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

#check proportion represented by each order, hymenoptera, diptera, coleoptera...
 tr<-read.csv("traits_ainhoa.csv")
# #head(tr)
# 
 d$order<-tr$order[match(d$Pollinator_gen_sp, tr$Pollinator_gen_sp)]
 d.order<- aggregate(Frequency ~ order, d, FUN= "sum")
 d.order$prop<-d.order$Frequency/sum(d.order$Frequency)

```

Within our sampling we recorded `r nrow(d)` plant-pollinator interactions involving `r nlevels(unique(d$Pollinator_gen_sp))` pollinator species and `r nlevels(unique(d$Plant_gen_sp))` plant species. Within the pollinator community the distribution of individuals in different orders was: `r round(d.order$prop[d.order$order=="Hymenoptera"]*100, digits=2)`% Hymenoptera, `r round(d.order$prop[d.order$order=="Diptera"]*100, digits=2)`% Diptera, `r round(d.order$prop[d.order$order=="Coleoptera"]*100, digits=2)`% Coleoptera and `r round(d.order$prop[d.order$order=="Lepidoptera"]*100, digits=2)`% Lepidoptera.

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

#######CALCULATE SAMPLING COMPLETENESS FOR POLLINATORS, PLANTS AND LINKS

#POLLINATORS
d2<-table(d$Site_ID, d$Pollinator_gen_sp)
d3<-as.data.frame.array(d2)

d4 <- t(d3[,1:ncol(d3)]) #I updated this to use defensive programming.
#head(d4)
#str(d4)
d5<-as.data.frame(d4)
#str(d5)
d6<-as.list(d5)

rar <- iNEXT(d6, q=0, datatype="abundance", endpoint = 400)
poll<-ggiNEXT.iNEXT(rar, color.var="site", se=FALSE) + ylab("Pollinator sps richness") + theme_bw() + 
  theme (legend.position= "none") + theme(text = element_text(size = 15), axis.text=element_text(size=rel(1)),
                                          axis.title.x = element_blank(),
        axis.title=element_text(size=rel(1),face="bold"), legend.direction="horizontal", legend.text=element_text(size=10)) +
  scale_color_manual(values=c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236", "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", "#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15", "#9C964A", "#CDC08C", "#FAD77B"))

#poll

#per site sampling completeness for pollinators
#Aznalcazar 64/103 = 0.62
#Bonares  54.000/178.701 = 0.30
#ConventodelaLuz  33.000/98.950 = 0.33
#CotitodeSantaTeresa  54.000/190.442 = 0.28
#Elpinar  29.000/158.657= 0.18
#Elpozo  32.000/174.000= 0.18
#Esparragal  39.000/110.859= 0.35
#LaCunya  32.000/118.672= 0.27
#LaRocina  74.000/147.129 = 0.50
#Lasmulas  52.000/127.295 = 0.41
#Niebla  57.000/124.792 = 0.46
#PinaresdeHinojos 40.000/168.852 = 0.24
#Pinodelcuervo  52.000/187.503 = 0.28
#Urbanizaciones  40.000/109.136 = 0.37 
#Villamanriqueeste  62.000/141.310 = 0.44
#Villamanriquesur  48.000/111.326 = 0.43

#(0.62 + 0.30 + 0.33 + 0.28 + 0.18 + 0.18 + 0.35 + 0.27 + 0.50 + 0.41 + 0.46 + 0.24 + 0.28 + 0.37 + 0.44 + 0.43)/16
#TOTAL AVERAGE sampling completeness of pollinators 0.35

#PLANTS
d.pl<-table(d$Site_ID, d$Plant_gen_sp)
d.pl2<-as.data.frame.array(d.pl)

d.pl3 <- t(d.pl2[,1:57])
#head(d.pl3)
#str(d.pl3)
d.pl4<-as.data.frame(d.pl3)
#str(d.pl4)
d.pl5<-as.list(d.pl4)

rar2 <- iNEXT(d.pl5, q=0, datatype="abundance", endpoint = 200)
plant<-ggiNEXT.iNEXT(rar2, color.var="site", se=FALSE) + ylab("Plant sps richness") + 
  theme_bw()  + theme (legend.position= "none") + theme(text = element_text(size = 15), axis.text=element_text(size=rel(1)),
                                                         axis.title.x = element_blank(),
        axis.title=element_text(size=rel(1),face="bold"), legend.direction="horizontal", legend.text=element_text(size=10)) +
  scale_color_manual(values=c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236", "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", "#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15", "#9C964A", "#CDC08C", "#FAD77B"))

#plant

#per site sampling completeness for plants
#Aznalcazar 17.000/18.490 = 0.92
#Bonares  13.000/13.990 = 0.93
#ConventodelaLuz  12.000/13.978 = 0.86
#CotitodeSantaTeresa  10.000/11.973 = 0.84
#Elpinar  11.000/23.255 = 0.47
#Elpozo  11.000 /13.219 = 0.83
#Esparragal  12.000 /16.436 = 0.73
#LaCunya  13.000/21.847 = 0.60
#LaRocina  16.000 /21.959 = 0.73 
#Lasmulas  12.000 /17.940 = 0.67
#Niebla  21.000/21.743 = 0.97
#PinaresdeHinojos 14.000/21.889 = 0.64
#Pinodelcuervo  15.000/27.372 = 0.55
#Urbanizaciones  10.000 /10.247 = 0.98
#Villamanriqueeste  9.000 /10.983 = 0.82
#Villamanriquesur  12.000 /12.989 = 0.92

#(0.92 + 0.93 + 0.86 + 0.84 + 0.47 + 0.83 + 0.73 + 0.60 + 0.73 + 0.67 + 0.97 + 0.64 + 0.55 + 0.98 + 0.82 + 0.92)/16

#TOTAL AVERAGE sampling completeness of plants 0.78

#LINKS

d$link<-paste(d$Plant_gen_sp, d$Pollinator_gen_sp)
d.l<-table(d$Site_ID, d$link)
d.l2<-as.data.frame.array(d.l)

d.l3 <- t(d.l2[,1:750])
#head(d.l3)
#str(d.l3)
d.l4<-as.data.frame(d.l3)
#str(d.l4)
d.l5<-as.list(d.l4)

rar3 <- iNEXT(d.l5, q=0, datatype="abundance", endpoint = 500)
link<-ggiNEXT.iNEXT(rar3, color.var="site", se=FALSE) + ylab("Link richness") + theme_bw() +
   theme (legend.position= "none") + theme(text = element_text(size = 15), axis.text=element_text(size=rel(1)),
        axis.title=element_text(size=rel(1),face="bold"), legend.direction="horizontal", legend.text=element_text(size=10)) +
 scale_color_manual(values=c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236", "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", "#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15", "#9C964A", "#CDC08C", "#FAD77B"))

#link

#per site sampling completeness for links
#Aznalcazar 101.000/244.145 = 0.41
#Bonares  72.000/389.031 = 0.19
#ConventodelaLuz  52.000/221.285 = 0.23
#CotitodeSantaTeresa  64.000/478.893 = 0.13
#Elpinar  37.000/194.026 = 0.19
#Elpozo   46.000/159.994 = 0.29
#Esparragal   52.000/196.900 = 0.26
#LaCunya  44.000/115.025 = 0.38
#LaRocina  98.000/351.931 = 0.28
#Lasmulas  69.000/306.883 = 0.22
#Niebla  84.000/611.776 = 0.14
#PinaresdeHinojos  56.000/186.413 = 0.30
#Pinodelcuervo   77.000/273.425 = 0.28
#Urbanizaciones  55.000/204.100 = 0.27
#Villamanriqueeste  76.000/193.208 = 0.39
#Villamanriquesur  67.000/234.221 = 0.29

#(0.41 + 0.19 + 0.23 + 0.13 + 0.19 + 0.29 + 0.26 + 0.38 + 0.28 + 0.22 + 0.14 + 0.30 + 0.28 + 0.27 + 0.39 + 0.29)/16

#TOTAL AVERAGE sampling completeness of links 0.27

#FIGURE S1 AT THE END OF THE PAPER

# grid.newpage()
# grid.draw(rbind(ggplotGrob(poll), ggplotGrob(plant), ggplotGrob(link), size = "last"))

```

  Our sampling completeness analyses revealed that with our survey we were able to capture 18-62% of pollinator species (average = 35%), 47-98% for plant species (average = 78%) and 13-41% for plant-pollinator links (average = 27%), in line with that obtained with other studies (e.g., [@Chacoff2012], Fig. S1). Our values of sampling completeness are slightly smaller in the case of pollinators, probably as a consequence of the great diversity found in the Mediterranean region and within our study area in particular, a hotspot of insect diversity [@Nieto2014a]. In addition, the fact that we include an extra effort to capture rare interactions observed outside of our main sampling might also increase the number of singletons which directly affect richness estimates.

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

# CALCULATE NICHE OVERLAP BETWEEN PLANT SPECIES AND TOTAL NUMBER OF VISITS AND JOIN FRUITSET DATA WITH NETWORK METRICS

#calculate total number of visits per plant sps

d.vis<-ddply(d, c("Site_ID", "Plant_gen_sp"), summarise,
            tot.visits= sum(Frequency))

d.vis$j<-paste(d.vis$Site_ID, d.vis$Plant_gen_sp)

#calculate average niche overlap per species

sites <- unique(d$Site_ID)

out.niche <- data.frame(Site_id = NA, row = NA, col = NA,value  = NA)

webs <- list()
for(i in 1:length(sites)){
  temp <- subset(d, Site_ID == sites[i])
  temp2<-droplevels(temp)

  web <- table(temp2$Pollinator_gen_sp, temp2$Plant_gen_sp)

  ni<-niche.overlap(web, method="morisita") #THIS SHOULD BE EXPLAINED IN METHODS
  df <- melt(as.matrix(ni), varnames = c("row", "col"))
  df$row<-as.character(df$row)
  df$col<-as.character(df$col)

  n <- nrow(out.niche)

  out.niche[n + seq(nrow(df)),1] <- as.character(sites[i])
  out.niche[n + seq(nrow(df)),2:3] <- (df[,1:2])
  out.niche[n + seq(nrow(df)),4] <-df[,3]

}

out.niche.agg <- ddply(out.niche, c("Site_id", "col"), summarise,
               mean.niche.over    = mean(value))

out.niche.agg$j<-paste(out.niche.agg$Site_id, out.niche.agg$col)

#use unaggregated data and join it with network metrics

f2$functional.comp.poll<-out.site$functional.comp.poll[match(f2$Site_ID, out.site$Site_id)]
f2$functional.comp.pl<-out.site$functional.comp.pl[match(f2$Site_ID, out.site$Site_id)]
f2$nodf.song<-out.site$nodf.song[match(f2$Site_ID, out.site$Site_id)]
f2$species.poll<-out.site$species.poll[match(f2$Site_ID, out.site$Site_id)]

#also match with plant species level network data
f2$match<-paste(f2$Site_ID, f2$Plant.sp)
outsp.pl.site$match<-paste(outsp.pl.site$Site_id, outsp.pl.site$species)

f2$norm_degree<-outsp.pl.site$norm_degree[match(f2$match, outsp.pl.site$match)]
f2$weigh_closeness<-outsp.pl.site$weigh_closeness[match(f2$match, outsp.pl.site$match)]

#and with niche overlap and total number of visits
f2$niche.overlap<-out.niche.agg$mean.niche.over[match(f2$match, out.niche.agg$j)]

f2$tot.visits<-d.vis$tot.visits[match(f2$match, d.vis$j)]
f2$tot.visits[is.na(f2$tot.visits)]<-0

#calculate number of pollinator species per plant species instead of norm_degree

sites <- unique(d$Site_ID)
plants<-unique(d$Plant_gen_sp)

out <- data.frame(Site_id = NA, plant.sps = NA, poll.sps = NA)

for(i in 1:length(sites)){
  temp <- subset(d, Site_ID == sites[i])
  for(j in 1:length(plants)){
      temp2 <- subset(temp, Plant_gen_sp == plants[j])
  temp3<-droplevels(temp2)

  poll<-nlevels(unique(temp3$Pollinator_gen_sp))

  n <- nrow(out)

  out[n + 1,1] <- as.character(sites[i])
  out[n + 1,2] <- as.character(plants[j])
  out[n + 1,3] <- poll

}
}
out

out2<-out[-1,]
out2$match<-paste(out2$Site_id, out2$plant.sps)

f2$pollspsperplant<-out2$poll.sps[match(f2$match, out2$match)]

#scale all variables used

f2$functional.comp.poll.sc<-scale(f2$functional.comp.poll, center = T, scale = T)
f2$functional.comp.pl.sc<-scale(f2$functional.comp.pl, center = T, scale = T)
f2$nodf.song.sc<-scale(f2$nodf.song, center = T, scale = T)
f2$species.poll.sc<-scale(f2$species.poll, center = T, scale = T)
f2$norm_degree.sc<-scale(f2$norm_degree, center = T, scale = T)
f2$weigh_closeness.sc<-scale(f2$weigh_closeness, center = T, scale = T)
f2$niche.overlap.sc<-scale(f2$niche.overlap, center = T, scale = T)
f2$pollspsperplant.sc<-scale(f2$pollspsperplant, center = T, scale = T)
f2$tot.visits.sc<-scale(f2$tot.visits, center = T, scale = T)

```

_Species-level analyses_

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#compare 2 models: one with simple diversity metrics (pollinator richness and number of visits), compared to one where position in a network (centrality) and niche overlap are included with these variables.

m1<-glmer(tot.fruitset ~  pollspsperplant.sc + tot.visits.sc  + (1|Plant.sp:Site_ID) + (1|Site_ID), family="binomial", data=f2)

summary(m1)
vif(m1)  #check for correlations between variables included in model

m2<-glmer(tot.fruitset ~  pollspsperplant.sc  + tot.visits.sc + weigh_closeness.sc + niche.overlap.sc  + (1|Plant.sp:Site_ID) + (1|Site_ID), family="binomial", data=f2)

summary(m2) 
vif(m2)

#select best model based on AIC

AICtab(m1, m2)

r2<-rsquared(c(m2, m1)) #check R^2 values for both models

# extract coefficients and p-values for m2 and present them in paper
coefs.m2 <- data.frame(coef(summary(m2)))
coefs.m2
Table1A<-round(coefs.m2[,1:3],digits=2)
rownames(Table1A)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Centrality", "Niche overlap")
#at the end of the paper

#plot significant variables, in this case centrality, do partial residual plots to show the effect of that variable in isolation of the others

#devtools::install_github("hohenstein/remef")  USE THIS PACKAGE TO DO PARTIAL RESIDUAL PLOTS OF SIGNIFICANT VARIABLES

y_partial_weigh <- remef(m2, fix= c( "tot.visits.sc", "pollspsperplant.sc", "niche.overlap.sc"), ran = "all", inverse = F)
#With this we remove the influece of x2 and the random effects from the dependent variable
y_partial_weigh_inv<-exp(y_partial_weigh)/(1+exp(y_partial_weigh))

Figure2a<-ggplot(m2@frame,  aes(x=m2@frame$weigh_closeness.sc, y=y_partial_weigh_inv, color= Plant.sp)) +    
  geom_point(size=2.5, shape=21, colour="black", aes(fill=factor(Plant.sp))) + stat_smooth(method=lm,  size=1, se=TRUE, color = "black", alpha=0.4) + 
  #geom_smooth(method=lm, linetype= "dashed", size=0.7, se=FALSE, alpha=0.8) +
  xlab("Centrality") +      
  ylab("Fruit set ") + theme_bw()  + theme (legend.position= "bottom") +
  theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1.3), face="bold")) + 
  scale_fill_discrete(name="Plant species") +  theme(legend.text = element_text(size = 14, face = "italic"), legend.title = element_text(size=15, face = "bold"), legend.direction = "vertical") + guides(fill=guide_legend(ncol=4))

#Figure2a

```

At the species level, in the case of fruit set, our results show that model 2 shows the best fit to our data (lowest AIC value), explaining `r round(r2$Conditional[1], 2)*100`% of the variability observed. In this case, we find a positive effect of a network structure metric, the centrality of a focal plant within the overall network on its fruit set (Table 1, Fig. 2A).

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

##seed set

s3$functional.comp.poll<-out.site$functional.comp.poll[match(s3$Site_ID, out.site$Site_id)]
s3$functional.comp.pl<-out.site$functional.comp.pl[match(s3$Site_ID, out.site$Site_id)]
s3$nodf.song<-out.site$nodf.song[match(s3$Site_ID, out.site$Site_id)]
s3$species.poll<-out.site$species.poll[match(s3$Site_ID, out.site$Site_id)]

#also match with plant species level network data
s3$match<-paste(s3$Site_ID, s3$Plant.sp)
outsp.pl.site$match<-paste(outsp.pl.site$Site_id, outsp.pl.site$species)

s3$norm_degree<-outsp.pl.site$norm_degree[match(s3$match, outsp.pl.site$match)]
s3$weigh_closeness<-outsp.pl.site$weigh_closeness[match(s3$match, outsp.pl.site$match)]
s3$niche.overlap<-out.niche.agg$mean.niche.over[match(s3$match, out.niche.agg$j)]

s3$tot.visits<-d.vis$tot.visits[match(s3$match, d.vis$j)]
s3$tot.visits[is.na(s3$tot.visits)]<-0

s3$pollspsperplant<-out2$poll.sps[match(s3$match, out2$match)]

#scale all variables used

s3$functional.comp.poll.sc<-scale(s3$functional.comp.poll, center = T, scale = T)
s3$functional.comp.pl.sc<-scale(s3$functional.comp.pl, center = T, scale = T)
s3$nodf.song.sc<-scale(s3$nodf.song, center = T, scale = T)
s3$species.poll.sc<-scale(s3$species.poll, center = T, scale = T)

s3$norm_degree.sc<-scale(s3$norm_degree, center = T, scale = T)
s3$weigh_closeness.sc<-scale(s3$weigh_closeness, center = T, scale = T)
s3$niche.overlap.sc<-scale(s3$niche.overlap, center = T, scale = T)
s3$pollspsperplant.sc<-scale(s3$pollspsperplant, center = T, scale = T)
s3$tot.visits.sc<-scale(s3$tot.visits, center = T, scale = T)

s3$mean.seedn.sc<-scale(s3$mean.seedn, center = T, scale = T)

```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}
#fit two models

m1.seed<-glmer(mean.seedn.sc ~  pollspsperplant.sc + tot.visits.sc  + (1|Plant.sp:Site_ID)+ (1|Site_ID), family="gaussian", data=s3)

summary(m1.seed)
vif(m1.seed)

m2.seed<-glmer(mean.seedn.sc ~  pollspsperplant.sc  + tot.visits.sc + weigh_closeness.sc + niche.overlap.sc  + (1|Plant.sp:Site_ID) + (1|Site_ID), family="gaussian", data=s3)

summary(m2.seed)
vif(m2.seed)

AICtab(m1.seed, m2.seed) #m2 is best

r2b<-rsquared(c(m2.seed, m1.seed)) #check R^2 values for both models

#test for problems in residuals
# res<-simulateResiduals(m2.seed)
# plot(res, rank = TRUE)

# extract coefficients and p-values for m1 and present them in paper
coefs.m2.seed <- data.frame(coef(summary(m2.seed)))
# use normal distribution to approximate p-value
#coefs.m2.seed$p.z <- 2 * (1 - pnorm(abs(coefs.m2.seed$t.value)))
coefs.m2.seed
Table1B<-round(coefs.m2.seed,digits=2)
rownames(Table1B)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Centrality", "Niche overlap")

#plot significant variables
y_partial_norm_seed <- remef(m2.seed, fix= c("pollspsperplant.sc", "tot.visits.sc", "weigh_closeness.sc"), ran = "all")
#With this we remove the influece of x2 and the random effects from the dependent variable

Figure2b<-ggplot(m2.seed@frame,  aes(x=m2.seed@frame$niche.overlap.sc, y=y_partial_norm_seed, color= Plant.sp)) +    
  geom_point(size=2.5, shape=21, colour="black", aes(fill=factor(Plant.sp))) + stat_smooth(method=lm,  size=1, se=TRUE, color = "black", alpha=0.4) + 
  #geom_smooth(method=lm, linetype= "dashed", size=0.7, se=FALSE, alpha=0.8) +
  xlab("Niche overlap") +      
  ylab("Number of \n seeds per fruit") + theme_bw()  + theme (legend.position= "right") +
  theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1.3), face="bold")) + 
  scale_fill_discrete(name="Plant species") +  theme(legend.text = element_text(size = 6, face = "italic"), legend.title = element_text(size=11, face = "bold"))

#Figure2b

#how to have two plots with same legend

prow <- plot_grid( Figure2a + theme(legend.position="none"),
                   Figure2b + theme(legend.position="none"),
                   align = 'vh',
                   labels = c("A", "B"),
                   hjust = -1,
                   nrow = 1
)

# extract the legend from one of the plots
legend_b <- get_legend(Figure2a + theme(legend.position="right"))

# add the legend underneath the row we made earlier. Give it 10% of the height
# of one plot (via rel_heights).
Figure2 <- plot_grid( prow, legend_b, ncol = 1, rel_heights = c(2, 1))#al final del paper

```

 For the average number of seeds per fruit, our results show again that model 2 shows the best fit, explaining `r round(r2b$Conditional[1], 2)*100`% of the variability observed in our data. In this case, we find a positive effect of the niche overlap between plant species on the number of seeds produced (Table 1B, Fig. 2B).

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#SEED WEIGHT

m1.seed.w<-glmmTMB(mean.seedw ~  pollspsperplant.sc + tot.visits.sc  + (1|Plant.sp:Site_ID), family="poisson", data=s3)

summary(m1.seed.w)

m2.seed.w<-glmmTMB(mean.seedw ~  pollspsperplant.sc + tot.visits.sc + weigh_closeness.sc + niche.overlap.sc  + (1|Plant.sp:Site_ID), family="poisson", data=s3)

summary(m2.seed.w)

#select best model based on AIC

AICtab(m1.seed.w, m2.seed.w)

# extract coefficients and p-values for m1 and present them in paper
tableS4<-round(summary(m1.seed.w)$coefficients$cond[,1:3], digits=2)
rownames(tableS4)<-c("(Intercept)", "Pollinator richness", "Total number of visits")

#FRUIT WEIGHT

m1.fruit.w<-glmmTMB(mean.fruitweight ~  pollspsperplant.sc + tot.visits.sc  + (1|Plant.sp:Site_ID) + (1|Site_ID), family="poisson", data=s3)

summary(m1.fruit.w)

m2.fruit.w<-glmmTMB(mean.fruitweight ~  pollspsperplant.sc  + tot.visits.sc + weigh_closeness.sc + niche.overlap.sc  + (1|Plant.sp:Site_ID) + (1|Site_ID), family="poisson", data=s3)

summary(m2.fruit.w)

#select best model based on AIC

AICtab(m1.fruit.w, m2.fruit.w)

# extract coefficients and p-values for m1 and present them in paper
tableS5<-round(summary(m2.fruit.w)$coefficients$cond[,1:3], digits=2)
rownames(tableS5)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Centrality", "Niche overlap")

```
 For all other measures of reproductive success considered (i.e., fruit and seed weight), both models had very similar fits, with ($\Delta AIC=2.2$) and 2, respectively. However, none of the variables included within our model explain the differences observed (Tables S4-S5).

_Site-level analyses_

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

######################## NTW STRUCTURE

#aggregate values at site level

f.ag <- ddply(f2, c("Site_ID", "Plant.sp", "nodf.song", "functional.comp.poll",
                    "functional.comp.pl", "species.poll"), 
               summarise, tot.visits    = sum(tot.visits),
                         mean.fruitset = mean(tot.fruitset))

f.agg <- ddply(f.ag, c("Site_ID",  "nodf.song", "functional.comp.poll", 
                       "functional.comp.pl","species.poll"), 
               summarise, tot.visits    = sum(tot.visits),
                         mean.fruitset = mean(mean.fruitset))

#scale all variables to be able to compare effect sizes

f.agg$nodf.song.sc <- scale(f.agg$nodf.song, center = T, scale = T)
f.agg$functional.comp.poll.sc <- scale(f.agg$functional.comp.poll, center = T, scale = T)
f.agg$functional.comp.pl.sc <- scale(f.agg$functional.comp.pl, center = T, scale = T)
f.agg$species.poll.sc <- scale(f.agg$species.poll, center = T, scale = T)
f.agg$tot.visits.sc <- scale(f.agg$tot.visits, center = T, scale = T)
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#FRUITset

m1.fruit.net<-glmmTMB(mean.fruitset ~  species.poll.sc + tot.visits.sc, family="beta_family", data=f.agg)

summary(m1.fruit.net)

m2.fruit.net<-glmmTMB(mean.fruitset ~  species.poll.sc + tot.visits.sc + nodf.song.sc + functional.comp.poll.sc, family="beta_family", data=f.agg)

summary(m2.fruit.net)

#removing one point with large number of visits to see if this is driving the effect of total number of visits (non-overlapping confidence intervals)

novis<-subset(f.agg, f.agg$tot.visits.sc<2)

 m2.fruit.novis<-glmmTMB(mean.fruitset ~  species.poll.sc + tot.visits.sc + nodf.song.sc + functional.comp.poll.sc, family="beta_family", data=novis)

summary(m2.fruit.novis) #still non-overlapping values

#select best model based on AIC

AICtab(m1.fruit.net, m2.fruit.net) #Almost as good. With this, you cannot say one is better than the other one. The take home message would be: if interested in predictions, go for simplicity. If interested in mechanisms, niche part is a mechanism to take into account. And actually m2 has better predictive ability (better R2 see below)

#provide R^2 values for both models

#crude alternative suggested by Ben Bolker to calculate R2 for glmmTMb models https://github.com/glmmTMB/glmmTMB/issues/169

r2.m2<-cor(f.agg$mean.fruitset,predict(m2.fruit.net,type="response"))^2
r2.m1<-cor(f.agg$mean.fruitset,predict(m1.fruit.net,type="response"))^2

#table with results for paper
Table2A<-round(summary(m2.fruit.net)$coefficients$cond, digits=2)
rownames(Table2A)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Nestedness", "Niche complementarity")
#at the end of paper

#MANUALLY CREATing PARTIAL RESIDUAL PLOTS

pred<-summary(m2.fruit.net)$coefficients$cond[1]+ (summary(m2.fruit.net)$coefficients$cond[2] * mean(f.agg$species.poll.sc)) + (summary(m2.fruit.net)$coefficients$cond[3] * mean(f.agg$tot.visits.sc)) + (summary(m2.fruit.net)$coefficients$cond[4]* mean(f.agg$nodf.song.sc)) + (summary(m2.fruit.net)$coefficients$cond[5] * f.agg$functional.comp.poll.sc)

res<-pred-f.agg$mean.fruitset
res.tr<-exp(res)/(1+exp(res)) #transform using logit back transformation

pl2 <- ggplot(f.agg,  aes(x=functional.comp.poll.sc, y= res.tr)) +    
        geom_point(size=2.5, shape=21, color="black", fill="darkgrey") + 
  stat_smooth(method = 'lm', size=1, se=TRUE, color = "black", alpha=0.4) +     
        ylim(0.3,0.9) +
        xlab("Niche complementarity") +      
        ylab("Fruit set") + theme_bw()  + #theme (legend.position= "none") +
        theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

pred2<-summary(m2.fruit.net)$coefficients$cond[1]+ (summary(m2.fruit.net)$coefficients$cond[2] * f.agg$species.poll.sc) + (summary(m2.fruit.net)$coefficients$cond[3] * mean(f.agg$tot.visits.sc)) + (summary(m2.fruit.net)$coefficients$cond[4]* mean(f.agg$nodf.song.sc)) + (summary(m2.fruit.net)$coefficients$cond[5] * mean(f.agg$functional.comp.poll.sc))

res2<-pred2-f.agg$mean.fruitset
res.tr2<-exp(res2)/(1+exp(res2))

pl3 <- ggplot(f.agg,  aes(x=species.poll.sc, y= res.tr2)) +    
        geom_point(size=2.5, shape=21, color="black", fill="darkgrey") + 
  stat_smooth(method = 'lm', size=1, se=TRUE, color = "black", alpha=0.4) +  
        ylim(0.3,0.9) +
        xlab("Pollinator richness") +      
        ylab("Fruit set") + theme_bw()  + #theme (legend.position= "none") +
        theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

pred3<-summary(m2.fruit.net)$coefficients$cond[1]+ (summary(m2.fruit.net)$coefficients$cond[2] * mean(f.agg$species.poll.sc)) + (summary(m2.fruit.net)$coefficients$cond[3] * f.agg$tot.visits.sc) + (summary(m2.fruit.net)$coefficients$cond[4]* mean(f.agg$nodf.song.sc)) + (summary(m2.fruit.net)$coefficients$cond[5] * mean(f.agg$functional.comp.poll.sc))

res3<-pred3-f.agg$mean.fruitset
res.tr3<-exp(res3)/(1+exp(res3))

pl4 <- ggplot(f.agg,  aes(x=tot.visits.sc, y= res.tr3)) +    
        geom_point(size=2.5, shape=21, color="black", fill="darkgrey") + 
  stat_smooth(method = 'lm', size=1, se=TRUE, color = "black", alpha=0.4) +
        ylim(0.3,0.9) +
        xlab("Total number of visits") +      
        ylab("Fruit set") + theme_bw()  + #theme (legend.position= "none") +
        theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#seed number

s.com <- ddply(s3, c("Site_ID", "Plant.sp", "nodf.song", "functional.comp.poll",
                     "functional.comp.pl", "species.poll"), 
               summarise, tot.visits    = sum(tot.visits),
                         mean.seedset = mean(mean.seedn),
                          mean.seedweight = mean(mean.seedw),
                          mean.fruitweight = mean(mean.fruitweight))

s.comm <- ddply(s.com, c("Site_ID", "nodf.song", "functional.comp.poll", 
                         "functional.comp.pl","species.poll"), 
               summarise, tot.visits    = sum(tot.visits),
                         mean.seedset = mean(mean.seedset),
                          mean.seedweight = mean(mean.seedweight),
                          mean.fruitweight = mean(mean.fruitweight))

#scale variables for comparison of effect sizes

s.comm$nodf.song.sc <- scale(s.comm$nodf.song, center = T, scale = T)
s.comm$functional.comp.poll.sc <- scale(s.comm$functional.comp.poll, center = T, scale = T)
s.comm$functional.comp.pl.sc <- scale(s.comm$functional.comp.pl, center = T, scale = T)
s.comm$species.poll.sc <- scale(s.comm$species.poll, center = T, scale = T)
s.comm$tot.visits.sc <- scale(s.comm$tot.visits, center = T, scale = T)

```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#NUMBER OF SEEDS PER FRUIT

m1.seed.net <- lm(mean.seedset ~  species.poll.sc + tot.visits.sc,
    data=s.comm,family="gaussian")

 summary(m1.seed.net)

m2.seed.net <- lm(mean.seedset ~  species.poll.sc + tot.visits.sc + nodf.song.sc + functional.comp.poll.sc, data=s.comm,family="gaussian")

 summary(m2.seed.net)

#select best model based on AIC 
AICtab(m1.seed.net, m2.seed.net)#both models are equally good

#test for problems in residuals
res<-simulateResiduals(m2.seed.net)
# plot(res, rank = TRUE)

#provide R^2 values for both models

#crude alternative suggested by Ben Bolker to calculate R2 for glmmTMb models https://github.com/glmmTMB/glmmTMB/issues/169

r2.m2.seed<-cor(s.comm$mean.seedset,predict(m2.seed.net,type="response"))^2
r2.m1.seed<-cor(s.comm$mean.seedset,predict(m1.seed.net,type="response"))^2

#save table with results for paper
Table2B<-round(summary(m2.seed.net)$coefficients[,1:3], digits=2)
rownames(Table2B)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Nestedness", "Niche complementarity") #end of paper

pl6<-ggplot(s.comm,  aes(x=functional.comp.poll.sc, y=mean.seedset))+      
        geom_point(size=2.5, shape=21, color="black", fill="darkgrey") + stat_smooth(method=lm,  size=1, se=TRUE, color = "black", alpha=0.4) + 
        #geom_smooth(method=lm, linetype= "dashed", size=0.7, se=FALSE, alpha=0.8) +
        xlab("Niche complementarity") +      
        ylab("Number of \n seeds per fruit") + theme_bw()  + #theme (legend.position= "none") +
        theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

#pl6

Figure3<-plot_grid(pl3, pl4, pl2, pl6, labels = c("A", "B", "C", "D"), nrow=2)  
```

At the site level, we find different patterns for fruit set and the number of seeds per fruit as compared to those for fruit and seed weight. In the case of fruit set and the number of seeds per fruit, we find that both model 1 and 2 are equally good in describing the differences observed (i.e.,$\Delta AIC<2$), and thus we are not able to choose one over the other [@Burnham2002]. This suggests model 2 is a good model despite its added complexity, and actually shows a better predictive ability than model 1 (R^2 = `r round(r2.m2, 2)` for model 2 versus `r round(r2.m1, 2)` for model 1 in the case of fruit set and R^2 = `r round(r2.m2.seed,2)` for model 2 versus `r round(r2.m1.seed, 2)` for model 1 in the case of the number of seeds per fruit) and therefore we will comment results for this model only. In particular, we find that both fruit set and the number of seeds per fruit are positively related to niche complementarity between pollinators (Tables 2, Fig. 3). Additionally, we find a negative effect of site-level pollinator richness and the total number of visits on average fruit set (Table 2A, Fig. 3). The number of visits for one of the sites was particularly large, yet this did not affect our results as removing this site from the analysis showed the same effect (Estimate $\pm$ SE = `r round(summary(m2.fruit.novis$sdr)[3], 2)` $\pm$ `r round(summary(m2.fruit.novis$sdr)[9], 2)`).

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#FRUIT WEIGHT

m1.fruitw.net<-glm(mean.fruitweight ~  species.poll.sc + tot.visits.sc, family="gaussian", data=s.comm)

summary(m1.fruitw.net)

 vif(m1.fruitw.net)

m2.fruitw.net<-glm(mean.fruitweight ~  species.poll.sc + tot.visits.sc + nodf.song.sc + functional.comp.poll.sc, family="gaussian", data=s.comm)

 summary(m2.fruitw.net)

 vif(m2.fruitw.net)

#select best model based on AIC

AICtab(m1.fruitw.net, m2.fruitw.net)

#test for problems in residuals
# res<-simulateResiduals(m1.fruitw.net)
# plot(res, rank = TRUE)

r2.fruitw1<-r.squaredGLMM(m1.fruitw.net)
r2.fruitw2<-r.squaredGLMM(m2.fruitw.net)

#save table with results for paper
tableS6<-round(summary(m1.fruitw.net)$coef[,1:3], digits=2)
rownames(tableS6)<-c("(Intercept)", "Pollinator species diversity", "Total number of visits")#end of paper

predtest<-summary(m1.fruitw.net)$coefficients[1]+ (summary(m1.fruitw.net)$coefficients[2] * s.comm$species.poll.sc) + (summary(m1.fruitw.net)$coefficients[3] * mean(s.comm$tot.visits.sc))

restest<-predtest-s.comm$mean.fruitweight

pltest <- ggplot(s.comm,  aes(x=species.poll.sc, y= restest)) +    
  geom_point(size=2.5, shape=21, color="black", fill="darkgrey") + 
  stat_smooth(method = 'lm', size=1, se=TRUE, color = "black", alpha=0.4) +
  #ylim(0.3,0.9) +
  xlab("Pollinator richness") +      
  ylab("Fruit weight (g)") + theme_bw()  + #theme (legend.position= "none") +
  theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

# crPlots(m1.fruitw.net, smooth=FALSE, terms = ~  species.poll.sc, main = "", ylab = "Mean fruit weight", col.lines= "black", pch=19, col= "black", bg = "darkgrey", cex=2, cex.axis=1.2, cex.lab=1.5, 
#         xlab = "Pollinator species diversity") #end of paper

```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#SEED WEIGHT

m1.seedw.net<-glm(mean.seedweight ~  species.poll.sc + tot.visits.sc, family="gaussian", data=s.comm)

summary(m1.seedw.net)

vif(m1.seedw.net)

m2.seedw.net<-glm(mean.seedweight ~  species.poll.sc + tot.visits.sc + nodf.song.sc + functional.comp.poll.sc, family="gaussian", data=s.comm)

summary(m2.seedw.net)

vif(m2.seedw.net)

#select best model based on AIC

AICtab(m1.seedw.net, m2.seedw.net) #m1 is better

#test for problems in residuals
# res<-simulateResiduals(m1.seedw.net)
# plot(res, rank = TRUE)

r2.seedw1<-r.squaredGLMM(m1.seedw.net)
r2.seedw2<-r.squaredGLMM(m2.seedw.net)

#save table with results for paper
tableS7<-round(summary(m1.seedw.net)$coef[,1:3], digits=2)
rownames(tableS7) <-c("(Intercept)", "Pollinator species diversity", "Total number of visits")#end of paper

predtest2<-summary(m1.seedw.net)$coefficients[1]+ (summary(m1.seedw.net)$coefficients[2] * s.comm$species.poll.sc) + (summary(m1.seedw.net)$coefficients[3] * mean(s.comm$tot.visits.sc))

restest2<-predtest2-s.comm$mean.seedweight

pltest2 <- ggplot(s.comm,  aes(x=species.poll.sc, y= restest2)) +    
  geom_point(size=2.5, shape=21, color="black", fill="darkgrey") + 
  stat_smooth(method = 'glm', size=1, se=TRUE, color = "black", alpha=0.4) +
  #ylim(0.3,0.9) +
  xlab("Pollinator richness") +      
  ylab("Seed weight (g)") + theme_bw()  + #theme (legend.position= "none") +
  theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

# crPlots(m1.seedw.net, smooth=FALSE,  main = "", ylab = "Mean seed weight", col.lines= "black", pch=19, col= "black", bg = "darkgrey", cex=2, cex.axis=1.2, cex.lab=1.5)

```

  Contrastingly, in the case of weight variables (fruit and seed weight), in both cases we find that the best model is model 1, i.e., that only including simple visitation metrics (R^2 = `r round(r2.fruitw1[2], 2)` in the case of fruit weight and `r round(r2.seedw1[2], 2)` in the case of seed weight). Here, we find a consistent positive effect of site-level pollinator richness for both weight descriptors (Tables S6-S7, Fig. 4).

_Equity in fruitset_

When evaluating the effect of differences in community composition and network structure for equity in reproductive success across the different species within a community we find that model 1 is the best model for all the thresholds considered (0.50, 0.25 and 0.75). Also, we find a significant negative relationship between equity and pollinator richness using the 0.75 threshold (Fig. 4C, Tables S8, S9, S10).

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#EQUITY SPECIES WITHIN COMMUNITIES
#using raw values of mean fruitset for each species at each site
#dataset f.ag
#number of species per site tableS1

#calculate number of sps per site producing more than 0.5 in fruitset

highprod<-subset(f.ag, f.ag$mean.fruitset>0.5)
highprod.sps<-ddply(highprod, "Site_ID", summarise, 
                   high.sps = length(Plant.sp))

#equity as proportion of high producing compared to total
highprod.sps$tot.sps<-tableS1$tot.sps[match(highprod.sps$Site_ID, tableS1$Site_ID)]
highprod.sps$equity<-highprod.sps$high.sps/highprod.sps$tot.sps

highprod.sps$nodf.song.sc<-f.agg$nodf.song.sc[match(highprod.sps$Site_ID, f.agg$Site_ID)]
highprod.sps$functional.comp.poll.sc<-f.agg$functional.comp.poll.sc[match(highprod.sps$Site_ID, f.agg$Site_ID)]
highprod.sps$species.poll.sc<-f.agg$species.poll.sc[match(highprod.sps$Site_ID, f.agg$Site_ID)]
highprod.sps$tot.visits.sc<-f.agg$tot.visits.sc[match(highprod.sps$Site_ID, f.agg$Site_ID)]
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#fit two models
 m1.eq.comm<-glm(equity ~  species.poll.sc + tot.visits.sc , family="binomial", data=highprod.sps)

 summary(m1.eq.comm)

 m2.eq.comm<-glm(equity ~  species.poll.sc + tot.visits.sc + nodf.song.sc + functional.comp.poll.sc , family="binomial", data=highprod.sps)
summary(m2.eq.comm)

#select best model based on AIC

AICtab(m1.eq.comm, m2.eq.comm) #m1 is best no variables significant

#save table with results for paper, both models are equally good so save both tables
tableS8<-round(summary(m1.eq.comm)$coef[,1:3], digits=2)
rownames(tableS8)<-c("(Intercept)", "Pollinator richness", "Total number of visits")

#test for problems in residuals
# res<-simulateResiduals(m1.eq.comm)
# plot(res, rank = TRUE)

```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#EQUITY SPECIES WITHIN COMMUNITIES 75% threshold
#using raw values of mean fruitset for each species at each site
#dataset f.ag
#number of species per site tableS1

#calculate number of sps per site producing more than 0.5 in fruitset

highprod75<-subset(f.ag, f.ag$mean.fruitset>0.75)
highprod75.sps<-ddply(highprod75, "Site_ID", summarise, 
                   high.sps = length(Plant.sp))

#equity as proportion of high producing compared to total
highprod75.sps$tot.sps<-tableS1$tot.sps[match(highprod75.sps$Site_ID, tableS1$Site_ID)]
highprod75.sps$equity<-highprod75.sps$high.sps/highprod75.sps$tot.sps

highprod75.sps$nodf.song.sc<-f.agg$nodf.song.sc[match(highprod75.sps$Site_ID, f.agg$Site_ID)]
highprod75.sps$functional.comp.poll.sc<-f.agg$functional.comp.poll.sc[match(highprod75.sps$Site_ID, f.agg$Site_ID)]
highprod75.sps$species.poll.sc<-f.agg$species.poll.sc[match(highprod75.sps$Site_ID, f.agg$Site_ID)]
highprod75.sps$tot.visits.sc<-f.agg$tot.visits.sc[match(highprod75.sps$Site_ID, f.agg$Site_ID)]
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#fit two models
 m1.eq.comm75<-glm(equity ~  species.poll.sc + tot.visits.sc , family="binomial", data=highprod75.sps)

 summary(m1.eq.comm75)

 m2.eq.comm75<-glm(equity ~  species.poll.sc + tot.visits.sc + nodf.song.sc + functional.comp.poll.sc , family="binomial", data=highprod75.sps)
summary(m2.eq.comm75)

#select best model based on AIC

AICtab(m1.eq.comm75, m2.eq.comm75) #m1 is best no variables significant

tableS9<-round(summary(m1.eq.comm75)$coef[,1:3], digits=2)
rownames(tableS9)<-c("(Intercept)", "Pollinator richness", "Total number of visits")

#partial residual plots

#test for problems in residuals
# res<-simulateResiduals(m1.eq.comm)
# plot(res, rank = TRUE)

pred.eq<-summary(m1.eq.comm)$coef[1] + (summary(m1.eq.comm)$coef[2]*highprod75.sps$species.poll.sc) + (summary(m1.eq.comm)$coef[3]*mean(highprod75.sps$tot.visits.sc))

res.eq<-pred.eq-highprod75.sps$equity
res.eq.retro<-exp(res.eq)/(1+exp(res.eq))

pl75 <- ggplot(highprod75.sps,  aes(x=species.poll.sc, y= res.eq.retro)) +    
        geom_point(size=2.5, shape=21, color="black", fill="darkgrey") + 
  stat_smooth(method = 'lm', size=1, se=TRUE, color = "black", alpha=0.4) +    
        xlab("Pollinator richness") +      
        ylab("Equity in fruitset") + theme_bw()  + #theme (legend.position= "none") +
        theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

Figure4<-plot_grid(pltest, pltest2, pl75, labels = c("A", "B", "C"), nrow=1)  
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#EQUITY SPECIES WITHIN COMMUNITIES 25% threshold
#using raw values of mean fruitset for each species at each site
#dataset f.ag
#number of species per site tableS1

#calculate number of sps per site producing more than 0.5 in fruitset

highprod25<-subset(f.ag, f.ag$mean.fruitset>0.25)
highprod25.sps<-ddply(highprod25, "Site_ID", summarise, 
                   high.sps = length(Plant.sp))

#equity as proportion of high producing compared to total
highprod25.sps$tot.sps<-tableS1$tot.sps[match(highprod25.sps$Site_ID, tableS1$Site_ID)]
highprod25.sps$equity<-highprod25.sps$high.sps/highprod25.sps$tot.sps

highprod25.sps$nodf.song.sc<-f.agg$nodf.song.sc[match(highprod25.sps$Site_ID, f.agg$Site_ID)]
highprod25.sps$functional.comp.poll.sc<-f.agg$functional.comp.poll.sc[match(highprod25.sps$Site_ID, f.agg$Site_ID)]
highprod25.sps$species.poll.sc<-f.agg$species.poll.sc[match(highprod25.sps$Site_ID, f.agg$Site_ID)]
highprod25.sps$tot.visits.sc<-f.agg$tot.visits.sc[match(highprod25.sps$Site_ID, f.agg$Site_ID)]
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}

#fit two models
 m1.eq.comm25<-glm(equity ~  species.poll.sc + tot.visits.sc , family="binomial", data=highprod25.sps)

 summary(m1.eq.comm25)

 m2.eq.comm25<-glm(equity ~  species.poll.sc + tot.visits.sc + nodf.song.sc + functional.comp.poll.sc , family="binomial", data=highprod25.sps)
summary(m2.eq.comm25)

#select best model based on AIC

AICtab(m1.eq.comm25, m2.eq.comm25) #m1 is best no variables significant

tableS10<-round(summary(m1.eq.comm25)$coef[,1:3], digits=2)
rownames(tableS10)<-c("(Intercept)", "Pollinator richness", "Total number of visits")

```

  Within our simulation evaluating the effect of niche complementarity on equity in reproductive success as more plants within the community are considered, we find that the effect of complementarity becomes more important as the reproductive success of more species is considered (Fig. 5). This importance seems to reach some sort of plateau at 6 species. However, this should be further evaluated, as this is the maximum number of species simultaneously observed in a community for our study, which precludes us from simulating further numbers of species.

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide', fig.show = 'hide'}

f2$Plant.sp<-paste(f2$Plant_genus, f2$Plant_species)
f3<-f2[,c(1,13:14)]

f4 <- ddply(f3, c("Site_ID", "Plant.sp"), summarise,
                         mean.fruitset    = mean(tot.fruitset))

table(f4$Plant.sp)
#Astragalus lusitanicus and Cistus albidus is causing problems later on. It has a very low fruit set, but as there is only one value, when correcting for max fruit set, gets a value of 1. Both are in Cotito... This is problematic. Similar with Anchusa azurea.

#I am removing them here.
f4 <- subset(f4, !Plant.sp %in% c("Astragalus lusitanicus", "Cistus albidus", "Anchusa azurea"))

#f4

fr.max<-aggregate(.~ Plant.sp, f4, FUN=max)

f4$fruit.max<-fr.max$mean.fruitset[match(f4$Plant.sp, fr.max$Plant.sp)]

f4$fruit.prop<-f4$mean.fruitset/f4$fruit.max
head(f4)

#1) First let's calculate all combinations of 1,2,3,4,5,6 species per network.
combinatorials_all <-  matrix(ncol = 5, nrow = 1, NA) 
colnames(combinatorials_all) <- c("SiteID", "Plant_number_level", "iteration", "equity", "functional.comp")
uniqueSite_ID <- unique(f4$Site_ID)
for(k in 1:length(uniqueSite_ID)){ #sites
  Site_temp <- subset(f4, Site_ID == uniqueSite_ID[k]) #1:SiteID
  n <- length(Site_temp$Plant.sp)
  combinatorials <- matrix(ncol = 5, nrow = 1, NA) #flexible
  colnames(combinatorials) <- c("SiteID", "Plant_number_level", "iteration", "equity", "functional.comp")
  combinatorials <- as.data.frame(combinatorials)
  for (i in 1:n){ #levels
    combinatorials_1 <- combn(x = Site_temp$Plant.sp, m = i) #1:n
    for (j in 1:ncol(combinatorials_1)){ #different combinations
      temp1 <- Site_temp[which(Site_temp$Plant.sp %in% combinatorials_1[,j]),]
      combinatorials_2 <- subset(temp1, temp1$fruit.prop > 0.5)
      if(dim(combinatorials_2)[1] == 0) {
        combinatorials_2 <- temp1
        combinatorials_2$no_rows <- 0
      } else{
        combinatorials_2$no_rows<-rep(1, nrow(combinatorials_2))
      }
      #head(combinatorials_2)
      #for each site here is the number of sps whose fruitset exceeds 50% of maximum fruitset observed in the whole study area
      combinatorials_3 <- aggregate(no_rows ~ Site_ID, combinatorials_2, sum)
      #head(combinatorials_3)
      combinatorials_3$functional.comp.poll <- out.site$functional.comp.poll[match(combinatorials_3$Site_ID, out.site$Site_id)]
      #now save for later.
      index <- dim(combinatorials)[1]
      combinatorials[index+1,1] <-  as.character(combinatorials_3$Site_ID) #1:SiteID
      combinatorials[index+1,2] <-  i #Plant number level
      #combinatorials[index+1,3] <-    #"iteration" let's asume we don' care, but would be good for unit testing
      combinatorials[index+1,4] <- combinatorials_3$no_rows/i #above threshold in %
      combinatorials[index+1,5] <-  combinatorials_3$functional.comp.poll #nich comp value...

    } #close different combinations
  } #close levels
  combinatorials_all <- rbind(combinatorials_all, combinatorials[-1,])
} #close sites

combinatorials_all <- combinatorials_all[-1,]

#plot overall

#scatter.smooth(combinatorials_all$equity ~ combinatorials_all$functional.comp)

#scale variables

combinatorials_all$functional.comp.sc<-scale(combinatorials_all$functional.comp, center = T, scale = T)

#now try overall, per level.
estimate <- rep(NA,6)
for (i in unique(combinatorials_all$Plant_number_level)){
  temp <- subset(combinatorials_all, Plant_number_level == i)
  m <- lm(temp$equity ~ temp$functional.comp.sc)
  plot(temp$equity ~ temp$functional.comp.sc) 
  abline(m)
  estimate[i] <- m$coefficients[2]
  }

#plot(unique(combinatorials_all$Plant_number_level) , estimate, las = 1, ylim = c(-0.01, 0.1) ,
   #  ylab = "importance of niche partitioning", xlab = "number of plants considered")

#now we subsample 16 combinations at a time 100 timeas and make the average.
estimate_i <- data.frame(level = NA, estimate = NA)
estimate_k <- data.frame(level = NA, estimate = NA)
head(combinatorials_all)
for (i in 1:length(unique(combinatorials_all$Plant_number_level))){ # over levels
  temp <- subset(combinatorials_all, Plant_number_level == unique(combinatorials_all$Plant_number_level)[i])
  for(k in 1:100){ #iterations
    value <- data.frame(site = NA, eq = NA, nich = NA)
    for(j in 1:length(unique(temp$SiteID))){ # over sites
      temp2 <- subset(temp, SiteID == unique(temp$SiteID)[j]) 
      value_temp <- temp2[sample(x = 1:nrow(temp2), 1),]
      value[j,1] <- value_temp$SiteID
      value[j,2] <- value_temp$equity
      value[j,3] <- value_temp$functional.comp
    } #close sites
    m <- lm(value$eq ~ value$nich)
    #plot(value$eq ~ value$nich)
    estimate_k[k,1] <- i
    estimate_k[k,2] <- m$coefficients[2]  
  } #close iteration
  #unique(estimate_k) #only 5 unique ombinations for n = 5. All negative. highly influenced by one data point of low niche part and equ = 1. "CotitodeSantaTeresa"
  #subset(f4, Site_ID == "CotitodeSantaTeresa") #Super high values!
  estimate_mean <- aggregate(estimate ~ level, data = estimate_k, FUN = "mean")
  estimate_i[i,1] <- estimate_mean$level
  estimate_i[i,2] <- estimate_mean$estimate
  }

estimate_i

#plot(estimate_i$level, estimate_i$estimate, las = 1,
    # ylab = "importance of niche partitioning", xlab = "number of plants considered")

Figure5<-ggplot(estimate_i,  aes(x=level, y=estimate)) +    
  geom_point(size=4) + 
  xlab("Number of plant species considered") +      
  ylab("Importance of niche complementarity") + theme_bw()  +
  theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

ggplot(s.comm,  aes(x=functional.comp.poll.sc, y=mean.seedset))+      
        geom_point(size=2.5, shape=21, color="black", fill="darkgrey") + stat_smooth(method=lm,  size=1, se=TRUE, color = "black", alpha=0.4) + 
        #geom_smooth(method=lm, linetype= "dashed", size=0.7, se=FALSE, alpha=0.8) +
        xlab("Niche complementarity") +      
        ylab("Number of seeds/fruit") + theme_bw()  + #theme (legend.position= "none") +
        theme(text = element_text(size = 18), axis.text=element_text(size=rel(1)), axis.title=element_text(size=rel(1), face="bold"))

```

**Discussion**

The existence of relationships between network structure and ecosystem function has been long hypothesized, yet, the specific mechanisms by which structure influences function have remained elusive until now [@Thompson2012]. Our results show that different aspects of network structure affect different dimensions of ecosystem functioning. In particular, we find that the centrality of a plant species within a community, which measures the number of connections it receives from other species in the community, has a positive effect for its fruit set. At the site level, we find that greater values of niche complementarity between pollinators result in larger average fruit sets.

  One of the first conclusions we can extract from the fact that in most cases both of the models we considered (i.e., the simple model based on visitation metrics and the more complex one including network structure metrics) were equally good, is that for predictive purposes simple visitation metrics, such as pollinator richness, might be enough [@Garibaldi2013; @Garibaldi2015]. Yet, a mechanistic understanding of the underlying processes driving these observed patterns requires that we take into account measures of network structure.

  Consistent with previous experimental [@Fontaine2005; @Frund2013], theoretical [@Pauw2013], and empirical studies [@Poisot2013; @Valdovinos2016], we find that niche complementarity is key in determining differences in reproductive outputs. Indeed, we find that communities where there is less overlap in the niches occupied by pollinator species had greater values of reproductive success, both greater fruit set and larger numbers of seeds per fruit. This therefore reflects the fact that reproductive success in plant species requires the delivery of conspecific pollen and thus of a certain degree of specialization amongst pollinator species on a particular plant resource in order to avoid the negative effects of inter-specific pollen deposition (e.g., pollen loss [@Flanagan2009] or interference with conspecific pollen [@Morales2008]).

  In our study, we did not find an effect of nestedness for reproductive success in any case. This metric, widely used across network analysis, and which is deemed to stabilize [@Bastolla2009] or have a null effect [@James2012] on natural communities, does not seem to play a direct role in ecosystem function measured as plant reproductive success. However, our study is limited to a maximum of six plant species per community, and including more species might reveal different patterns, in which nestedness and the redundancy it implies might play a more important role.

  Site-level plant reproductive success, measured as average fruit or seed set across all the species considered is an important part of the functions delivered by pollinators to plants. However, these average values might be masking a great deal of variability amongst plant species, and thus a nuanced view of the effect of pollinators on whole-plant ensembles is needed. This can be captured by the effect of pollinators on equity in reproductive success across plant species. This aspect ensures that reproductive success is equally distributed amongst a larger number of species, thus contributing to the maintenance of greater species diversity values in natural populations. Indeed, we know that plant species diversity within a community is largely driven by different types of direct and indirect interactions including those amongst plant species (e.g., resource competition [@Goldberg1992] or facilitation [@Bruno2003]), as well as those defining antagonistic (e.g., involving pathogens [@Bagchi2010], or mutualistic interactions (e.g, pollinators [@Benadi2013; @Lanuza2018]). However, equitability in reproductive success across species is seldom taken into account, despite its importance in maintaining genetic diversity and ensuring the resilience of populations to further change.

  In the case of equity, we did not find a significant effect of either simple visitation or network structure metrics. However, the results of our simulation on the importance of network structure as the number of plant species considered increases, shows us that this effect dramatically increases when more than four plant species are considered. This implies that if we were able to measure reproductive success for all the plant species in all the communities (which is not feasible given constraints in sampling effort), we might find that the effects of network structure on equity might be even more prevalent.

  One of the unexpected results of our analyses is the strong negative effect of pollinator richness for fruit set at the site level, as well as a negative effect on equity in reproductive success at the 0.75 threshold. An explanation to this might be the fact that pollinator richness here includes all the pollinators recorded during our sampling efforts, i.e., it includes species that do not pollinate some of the species whose reproductive success was measured. More complex communities (with more pollinators, but also with more plant species, with a Pearson correlation value of 0.42 in our case) may require stabilizing mechanisms that reduce the competition exerted by the dominant plant species. This can be reached, for example, by a reduction in the reproductive success of the dominant species [@Lanuza2018], which are precisely those evaluated in this study. These ideas open the door to explore the positive or negative effects of the complete pollinator community on full plant species coexistence, which may be determined by density-dependence effects [@Benadi2018]. In our case, while fruit set is negatively related to pollinator richness, fruit and seed weight show the opposite relationship, indicating that this density-dependent effect might only be limiting fruit quantity and not fruit quality. Thus, taking into account the co-flowering plant species densities may be the next step [@Vanbergen2014].

  In summary, our findings show that the analysis of natural communities of interacting species using network analysis not only represents an ideal way of visualizing and grasping the complexity present within these communities, but also represents a manner of mechanistically understanding differences observed across the reproductive success of individuals and/or species. Future studies could build on this body of research by focusing on the long-lasting effects of network structure for plant fitness, by evaluating the consequences of different network structures on other stages of a plant´slife cycle (e.g., seed viability) and whether these differences are passed on to later generations (e.g., though different survival rates).

**Acknowledgements**

The authors would like to thank Oscar Aguado for help in identifying pollinator species. AM received funding from a Juan de la Cierva (IJCI-2014-22558) and Ikerbasque fellowships. IB acknowledges funding from MSC-PCIG14-GA-2013-631653 BeeFun Project. We thank Doñana’s Singular Scientific-Technical Infrastructure (ICTS-RBD) for access to the park.

\nextpage

##Tables

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5}
# extract coefficients and p-values for m2 and present them in paper
# coefs.m2 <- data.frame(coef(summary(m2)))
# Table1A<-round(coefs.m2[,1:3],digits=2)
# rownames(Table1A)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Centrality", "Niche overlap")

kable(Table1A, caption = "Table 1A. Results of GLMM showing effect of simple visitation and network structure metrics on species-level fruitset based on best model selected. Bold letters indicate significant variables.")
```
\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5}
# # extract coefficients and p-values for m1 and present them in paper
# coefs.m2.seed <- data.frame(coef(summary(m2.seed)))
# # use normal distribution to approximate p-value
# #coefs.m2.seed$p.z <- 2 * (1 - pnorm(abs(coefs.m2.seed$t.value)))

# Table1B<-round(coefs.m2.seed,digits=2)
# rownames(Table1B)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Centrality", "Niche overlap")

kable(Table1B, caption = "Table 1B. Results of GLMM showing effect of simple visitation and network structure metrics on species-level number of seeds/fruit based on best model selected. Bold letters indicate significant variables.")
```
\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5}
# #table with results for paper
# Table2A<-round(summary(m2.fruit.net)$coefficients$cond[,1:3], digits=2)
# rownames(Table2A)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Nestedness", "Niche complementarity")

kable(Table2A, caption = "Table 2A. Results of GLM showing effect of simple visitation and network structure metrics on site-level average fruit set based on best model selected.")

```
\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5}
# #save table with results for paper
# Table2B<-round(summary(m2.seed.net)$coefficients[,1:3], digits=2)
# rownames(Table2B)<-c("(Intercept)", "Pollinator richness", "Total number of visits", "Nestedness", "Niche complementarity")

kable(Table2B, caption = "Table 2B. Results of GLM showing effect of simple visitation and network structure metrics on site-level average number of seeds per fruit based on best model selected.")
```
\newpage

**Figure legends**

**Figure 1.** Map showing location of 16 study sites. Inset shows location of study area within SW Spain.

**Figure 2**. Partial residual plot showing the effect of a single predictor for the relationship between A) plant species centrality and fruit set for each of the plant species considered and B) plant niche overlap and average number of seeds per fruit. Dots represent each of the individuals sampled for each species within each site.

**Figure 3.** Partial residual plots showing the effect of A) pollinator richness, B) total number of visits received and C) niche complementarity between pollinator species on site-level average fruitset and D) niche complemenarity between pollinator species on the average number of seeds per fruit at the site level. Dots represent average values of fruit set at the level of the site for all plant species considered (N=16 sites).

**Figure 4.** Partial residual plots showing the effect of pollinator richness on site-level average A) fruit and B) seed weight and C) equity in reproductive success across species within a site (75% threshold). Dots represent values for each site (N=16 sites).

**Figure 5.** Results of simulation evaluating the importance of niche complementarity in determining differences in equity in reproductive across communities harboring from one to six species. Points represent average values across 1,000 simulated combinations.

\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=7, fig.width=10}

#FIGURE 1

mydf <- read.csv("site_coordinates.csv")

# Making a data frame *spatial*
library(devtools)
# install_github("edzer/sfr")
# install_github("edzer/rgdal2")
library(rgdal)
library(sf)
library(sp)
library(raster)
library(tmap)
library(dplyr)
library(spData)
library(leaflet)
library(ggplot2)
library(grid)

#devtools::install_github("Nowosad/spDataLarge")
library(spDataLarge)

occs <- st_as_sf(mydf, coords = c("longitude", "latitude"))

#setting projection
st_crs(occs) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +init=epsg:3042"

data(world)
sp<-subset(world, world$iso_a2==c("ES", "PT"))

#extent(occs)  #to calculate xmin, xmax, ymin, ymax
doñana_region = st_bbox(c(xmin = -7.000000, xmax = -6.00000,
                      ymin = 37.10000, ymax = 37.51200),
                    crs = st_crs(sp)) %>% 
  st_as_sfc()

sp_map = tm_shape(sp) + tm_polygons() +   tm_compass(type = "8star", position = c("right", "bottom")) +
  tm_shape(occs) + tm_symbols(shape = 16, col = "black", size = 0.03) + 
 tm_shape(doñana_region) + tm_borders(lwd = 3) #map of all spain

# Get elevation data from the internet 
elev <- getData("alt", country = "Spain")

# Crop raster to desired extent
ele <- crop(elev, c(-7.000000, -6.00000, 37.10000, 37.51200))

sp_points_map = tm_shape(ele, bbox = doñana_region) +
 tm_raster(legend.show = TRUE, title = "Elevation \n (m)") + 
  tm_layout(legend.text.size=1.5, legend.title.size = 1.6, legend.bg.color = "white", legend.frame = TRUE) +
  tm_shape(occs) + tm_symbols(shape = 24, col = "darkgrey", size = 2) +
  tm_scale_bar(position = c("left", "bottom"), size=1.5)

sp_points_map
print(sp_map, vp = viewport(0.8335, 0.6812, width = 0.35, height = 0.35))

```
**Figure 1**

\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=6, fig.width=9}

#FIGURE 2
Figure2

   # save_plot("figures/Figure2.tiff", Figure2,
   #           base_height = 6, base_width = 9)
```
**Figure 2**

\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=3, fig.width=11}

#FIGURE 3
Figure3

    # save_plot("figures/Figure3.tiff", Figure3,
    #           base_height = 3, base_width = 11)
```
**Figure 3**

\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=3, fig.width=10}

#FIGURE 4
Figure4
# 
# save_plot("figures/Figure4.tiff", Figure4,
#              base_height = 5, base_width = 10)
```
**Figure 4**

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5, fig.width=8}
Figure5
```
**Figure 5.**

\newpage

#Supplementary material

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=13, fig.width=7}

#######FIG S1 SAMPLING COMPLETENESS

figures1<-plot_grid( poll, plant, link, ncol = 1)
figures1

# save_plot("figures/FigureS1.tiff", figures1,
#            base_height = 13, base_width = 7)
```

**Figure S1.** Accumulation curves of pollinator, plant and plant-pollinator link richness with increasing sampling effort up to 100% sample coverage. Solid lines and points indicate observed richness while dashed lines show expected richness at increasing sample size, (i.e., extrapolated).

\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=3, fig.width=4}

crPlots(m1.fruitw.net, smooth=FALSE, terms = ~  species.poll.sc, main = "", ylab = "Mean fruit weight", col.lines= "black", pch=19, col= "black", bg = "darkgrey", cex=2, cex.axis=1.2, cex.lab=1.5, 
        xlab = "Pollinator species diversity") #end of paper
```

**Figure S2.** Partial residual plots showing the effect of the total number of pollinator visits received at the site level on the average fruit weight. Dots represent average vaues at the level of the site for all plant species considered (N=16 sites).

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5, fig.width=8}
crPlots(m1.seedw.net, smooth=FALSE,  main = "", ylab = "Mean seed weight", col.lines= "black", pch=19, col= "black", bg = "darkgrey", cex=2, cex.axis=1.2, cex.lab=1.5)
```

**Figure S3.** Partial residual plots showing the effect of the the diversity of pollinators and the total number of pollinator visits received at the site level on the average seed weight. Dots represent average vaues at the level of the site for all plant species considered (N=16 sites).

\newpage

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

#calculate number of plant species per site, number of individuals per plant species and number of fruits per ind surveyed for repr succ

f.inds<-ddply(f, c("Site_ID", "Plant.sp"), summarise, 
                   tot.inds = length(Plant.sp)) #number of individuals per plant sps

tableS1<-ddply(f.inds, c("Site_ID"), summarise, 
                   tot.sps = length(Site_ID)) #number of plant sps per site

colnames(tableS1)<-c("Site", "Number of plant species")

kable(tableS1, caption = "Table S1. Number of species sampled at each site.")

tableS2 <- aggregate(Plant.ID ~ Site_ID + Plant.sp, data = f2, FUN = function(x){NROW(x)})

colnames(tableS2)<-c("Site", "Plant species", "Number of individuals")

kable(tableS2, caption = "Table S2. Number of individuals per plant species sampled at each site.")

tableS3<-ddply(s, c("Site_ID", "Plant.sp"), summarise, 
                   tot.fruits = length(Plant.ID)) #number of fruits per individual

colnames(tableS3)<- c("Site", "Plant species", "Number of fruits")

kable(tableS3, caption = "Table S3. Number of fruits per plant species sampled at each site.")
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5}

tableS4<-round(summary(m1.seed.w)$coefficients$cond[,1:3], digits=2)
rownames(tableS4)<-c("(Intercept)", "Pollinator richness", "Total number of visits")
kable(tableS4, caption="Table S4. Results of GLMM showing effect of simple visitation metrics on species-level average seed weight based on best model selected.")

tableS5<-round(summary(m1.fruit.w)$coefficients$cond[,1:3], digits=2)
rownames(tableS5)<-c("(Intercept)", "Pollinator richness", "Total number of visits")
kable(tableS5, caption="Table S5. Results of GLMM showing effect of simple visitation metrics on species-level average fruit weight based on best model selected.")
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5}

tableS6<-round(summary(m1.fruitw.net)$coef[,1:3], digits=2)
rownames(tableS6)<-c("(Intercept)", "Pollinator richness", "Total number of visits")

kable(tableS6, caption="Table S6. Results of GLM showing effect of simple visitation metrics on site-level average fruit weight based on best model selected.")
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5}
#save table with results for paper
tableS7<-round(summary(m1.seedw.net)$coef[,1:3], digits=2)
rownames(tableS7) <-c("(Intercept)", "Pollinator richness", "Total number of visits")

kable(tableS7, caption="Table S7. Results of GLM showing effect of simple visitation metrics on site-level average seed weight based on best model selected.")
```

```{r, include=TRUE, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.height=5}

kable(tableS8, caption="Table S8. Results of GLM showing effect of simple visitation metrics on equity in reproductive success across plant species within a site based on best model selected (0.50 threshold).")

kable(tableS9, caption="Table S9. Results of GLM showing effect of simple visitation metrics on equity in reproductive success across plant species within a site based on best model selected (0.75 threshold).")

kable(tableS10, caption="Table S10. Results of GLM showing effect of simple visitation metrics on equity in reproductive success across plant species within a site based on best model selected (0.25 threshold).")
```

\newpage

**References**